Learning Excursion Sets of Vector-valued Gaussian Random Fields for Autonomous Ocean Sampling

### 2021-01-22 ans_to_reviewers/ans-to-2nd-rev.tex
Updates from Overleaf

We are grateful to the editors and reviewers for their comments and suggestions in this 2$^{nd}$ review. They have all been considered in this revised version of our manuscript. The changes in the manuscript are reflected in the points shown below.\\

@@

\reply: We agree and have addressed this point in \aecom \textbf{1}.

@@

\vspace{5mm}
\reply: We have elaborated on the specification of the nugget effect in the revised version. In the revised version we include a plot with analysis of survey sampling residuals at the end of the paper. This shows that the specified model is reasonable, but it is of course a simplification of the complex ocean. We believe this part fits as an honest discusion at the end of the results section. We have addressed this point further in \aecom \textbf{5}.

@@

\vspace{5mm}
\reply: We agree on all our points here, except Fig.8. Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c. So we ask to consider that this figure is kept in color. Revised Fig. 3 and 11 are in black-and-white. Fig. 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.

@@

\vspace{5mm}
\reply: We have elaborated on the specification of nugget in the revised version. The initial sampling transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect. The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, and in related data-sets. However, based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time. This variability is partly due to dynamic water masses but also uncertainty in the AUV position. 
\begin{figure}[h] 
\centering
\includegraphics[width=0.6\textwidth]{Figures/field-trials/empVAR.png}
\caption{Empirical variogram of the scaled variables from the two runs. }
\label{empVar}
\end{figure}

In the revised manuscript, we first clarified the specification of model parameters, and we then made residual plots and empirical variograms for the actual survey data (Figure 11 in the revised paper). In the two surveys, we have some samples close in space and it is possible to extract the nugget effect. These curves support our statements of a small but non-zero nugget.

@@

We further checked the sensitivity of the nugget specification in the sequential sampling algorithm. This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels. We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
Indeed, with small nugget, the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path. 
We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the excursion set. %We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if the sampling continues (here stopped for presentation purposes).

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.52]{ans_to_reviewers/merged_paths_2.png}
    \caption{Comparison of sampling paths and posterior EPs for different values of the measurement noise standard deviation (from left to right: $0.1$ and $0.5$). The sampling has been stopped after 200 steps.}
    \label{fig:nugget_comparison}
\end{figure}
In the revised manuscript we wrote another sentence on Fig 11: 'The AUV paths in both runs are relatively
winding in the sense that the vehicle reacts to the data. In our
experience from simulations, the paths tend to be smoother when we
use larger measurement error variances for the AUV temperature and salinity observations. '

>>>>

We are grateful to the editors and reviewers for their comments and suggestions. They have all been considered in this revised version of our manuscript. The changes in the manuscript are reflected in the points shown below.\\

@@

\reply: We agree, and we have addressed this point as you suggest. For more details, see \aecom \textbf{1}.

@@

\vspace{5mm}
\reply: We have revised the text specification of model parameters. In the revised version we have a plot with analysis of survey sampling residuals at the end of the paper. This shows that the specified nugget effect and other parameters are reasonable, but it is of course a simplification of the complex ocean. We believe this part fits as an honest discussion at the end of the results section. We have addressed this point further in \aecom \textbf{5}.

@@

\vspace{5mm}
\reply: We agree on all your points here, except Fig.8. Line type will be tricky for Fig. 8, when lines are very close together especially in Fig.8b, and Fig.8c. So we ask to consider that this figure is kept in color. Revised Fig. 3 and 11 are in black-and-white. Fig. 4, 5, and 6 have been made with a black-and-white compatible color-map, so they can be converted to black-and-white.

@@

\vspace{5mm}
\reply: We have revised the text specification of model parameters in Sect. 4.1 and 4.2. The initial survey transect had no samples at very short distance, and these data are hence not reliable for extracting the nugget effect. The sensor-provided standard deviations of the measurement tools are in the fourth decimal points, but based on our experience \citep{fossuminformation}, there is more variability than this in temperature and salinity measurements in water samples close in space and time. This variability is partly due to dynamic water masses but also uncertainty in the AUV position. 
\begin{figure}[h] 
\centering
\includegraphics[width=0.6\textwidth]{Figures/field-trials/empVAR.png}
\caption{Empirical variogram of the scaled variables from the two runs. }
\label{empVar}
\end{figure}

In the revised manuscript, we made residual plots and empirical variograms for the actual survey data in Sect 4.3 (Figure 11 in the revised paper). In the two surveys, we have some samples close in space and it is possible to extract the nugget effect. These curves support our statements of a small but non-zero nugget (see Figure \ref{empVar}).

@@

We further checked the sensitivity of the nugget specification in the sequential sampling algorithm. This was done by running synthetic cases of sampling paths and data up to a certain stage for different measurement noise variance levels. We see moderate changes in the sampling path patterns, but paths tend to be smoother for large nugget effect, as illustrated in Figure \ref{fig:nugget_comparison}.
Indeed, with small nugget (left display), the AUV reacts more to the observed temperature and salinity levels, and this leads to more small turns in its path.

We note that even though the sampling paths can be different, the posterior EPs in Figure \ref{fig:nugget_comparison} are similar and will thus lead to similar estimates for the ES. %We also remark that the differences in Figure \ref{fig:nugget_comparison} would get smaller if the sampling continues (here stopped for presentation purposes).

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.52]{ans_to_reviewers/merged_paths_2.png}
    \caption{Comparison of sampling paths and posterior EPs for different values of the measurement noise standard deviation (from left to right: $0.1$ and $0.5$). The sampling has been stopped after 200 steps.}
    \label{fig:nugget_comparison}
\end{figure}
In the revised manuscript we wrote two sentences around this sensitivity near Fig 10: 'The AUV paths in both runs are relatively
winding in the sense that the vehicle reacts to the data. In our
experience from simulations, the paths tend to be smoother when we
use larger measurement error variances for the AUV temperature and salinity observations. '

### 2021-01-22 main_web.tex
Updates from Overleaf

\address{\\Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e2}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e4}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e5}}

@@

\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
  \hfill
  \subfigure[Quadratic form CDFs.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[width = 0.32\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
  \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Fig.~\ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity
and temperature are subtracted from the data. This scatter-plot of
joint residuals indicates that the correlation between the two
variables might be slightly larger than the initially specified $0.5$ correlation,
and that initial uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
and theoretical curves are not very different, which indicates that
the specified Gaussian model fits reasonably
well. Fig.~\ref{fig:parest_c} shows the empirical variograms of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m as was
specified. The working assumption of a separable covariance function
is hence not unreasonable. There appears to be a slight nugget effect, but not much different from what was set as the measurement noise variance.
Overall, the spatial Gaussian model representation is a simplification
of the changing water conditions, and even within our working assumptions,
the parameter specification could likely have been
improved. However, we do not expect that minor changes in parameters would have caused much
difference in AUV sampling paths.

>>>>

\address{Trygve Olav Fossum \\Department of Marine Technology\\ Otto Nielsens veg. 10, 7491 Trondheim\\ Norway\\
\printead{e1}}
\address{Cédric Travelletti\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e2}}
\address{Jo Eidsvik\\Department of Mathematical Sciences\\ Hogskoleringen 1, 7491 Trondheim\\ Norway\\ \printead{e3}}
\address{David Ginsbourger\\ Institute of Mathematical Statistics and Actuarial Science \\ University of Bern \\
Switzerland.
\printead{e4}}
\address{Kanna Rajan\\Underwater Systems and Technology Laboratory,
  Faculty of Engineering,\\ Rua Dr. Roberto Frias\\ University of Porto, Portugal\\
\printead{e5}}

@@

\begin{figure}[!b]
  \centering
  \subfigure[Residual plot.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_aNEW.eps}\label{fig:parest_a}}
  \hfill
  \subfigure[Quadratic form CDFs.]{\includegraphics[height = 0.26\textwidth]{Figures/res_diag_bNEW.eps}\label{fig:parest_b}}
  \hfill
  \subfigure[Empirical variogram.]{\includegraphics[height = 0.252\textwidth]{Figures/res_diag_cNEW.eps}\label{fig:parest_c}}
  \caption{Data analysis from AUV experiments. \ref{fig:parest_a} Residual plot of temperature and salinity
    along with specified Gaussian contours. \ref{fig:parest_b} Empirical CDF
    (solid) of the quadratic form of the residuals along with the
    modeled theoretical CDF (dashed) of the $\chi^2$ distribution with two
    degrees of freedom. \ref{fig:parest_c} Scaled empirical variograms of the survey
    salinity and temperature data.}
\label{fig:parest}
\end{figure}

Fig.~\ref{fig:parest} summarizes with diagnostic plots of this
analysis. Fig.~\ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the regression mean values of salinity
and temperature are subtracted from the data. This scatter-plot of
joint residuals indicates that the correlation between the two
variables might be slightly larger than the initially specified $0.5$ correlation,
and that initial uncertainty in temperature was set a bit too high. Based on
the fitted bivariate covariance model (ellipses in
Fig.~\ref{fig:parest_a}), we can compute the scalar quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig.~\ref{fig:parest_b} shows
the empirical CDF of the quadratic forms (solid) together with the
theoretical CDF of the $\chi^2_2$ distribution (dashed). The empirical
and theoretical curves are not very different, which indicates that
the specified Gaussian model fits reasonably
well. Fig.~\ref{fig:parest_c} shows the empirical variograms of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m as was
specified. The working assumption of a separable covariance function
is hence not unreasonable. There appears to be a slight nugget effect, but not much different from what was set as the measurement noise variance.
Overall, the spatial Gaussian model representation is a simplification
of the changing water conditions, and even within our working assumptions,
the parameter specification could likely have been
improved. However, we do not expect that minor changes in parameters would have caused much
difference in AUV sampling paths.
